---
title: "Final Project: Visualizations"
author: "Thomas Bahng"
date: "March 9, 2020"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
library(data.table)
library(ggplot2)
library(knitr)
library(dplyr)
library(viridis)
library(purrr)
library(reshape2)
opts_chunk$set(echo = TRUE)
```

* This file creates some visualizations to support modeling and presentation.

# Plot Trends in Listings and Price Over Time
* Annual listing data for New York City was collected from [insideairbnb](http://insideairbnb.com/get-the-data.html) on March 9, 2020.
* Each of the five files are exclusive listings datasets from 2015 - 2019

### Read Yearly Listing Summary Data
* December (year-end) listings data collected for each year from 2015-2019.

```{r}
# read yearly listing summary data
fnames = list.files('data', pattern = 'listings_201\\d12.csv', full.names = TRUE)
df_list = lapply(
  fnames,
  function(x) fread(x)
)
# name each data set
names(df_list) <- 2015:2019
# dimensions of each data set
sapply(df_list, dim)
```

### Plot: Growth in Listings Over Time

```{r}
df <- map_df(df_list, ~as.data.frame(.x), .id="year")
dfplot <- df %>% count(year, neighbourhood_group)
head(dfplot) %>% kable(., row.names = FALSE)
```

```{r}
nycGrowth <- data.frame(
  year = 2016:2019,
  perc_growth = round(diff(tapply(dfplot$n, dfplot$year, sum)) / tapply(dfplot$n, dfplot$year, sum)[1:4],2)
)
```

```{r}
ggplot(nycGrowth, aes( x = year, y = perc_growth )) +
  geom_point() + geom_line() +
  ylab("percent change in listings") +
  ggtitle("Rate of Change in Number of Listings in NYC (2015 - 2019)")
```

*The growth rate in AirBnb listings in NYC has significantly decreased from 2015 - 2019.*

```{r}
ggplot(dfplot, aes(x = year, y = n, fill = neighbourhood_group)) +
  geom_col() +
  scale_fill_viridis(discrete = TRUE) +
  ylab("Count of Listings") +
  ggtitle("Total Number of AirBnb Listings in NYC by Year",
          subtitle = "Segmented by Neighbourhood Group")
```

*Overall trend shows an increase in listings, with the majority share of listings in Brooklyn and Manhattan.*

```{r}
# listings growth rate by neighbourhood
ngroups <- c('Bronx','Brooklyn','Manhattan','Queens','Staten Island')
# function to get YoY growth rate given neighbourhood group
getGrowth <- function(x) {
  d <- dfplot[dfplot$neighbourhood_group == x,]
  arr <- tapply(d$n, d$year, sum)
  out <- round(diff(arr)/arr[1:length(arr)-1], 2)
  return(out)
}
# dataframe of YoY change
dYearDelta <- as.data.frame(sapply(ngroups, getGrowth) %>% t)
dYearDelta$neighbour_group <- ngroups
dYearDelta <- dYearDelta[,c('neighbour_group','2016','2017','2018','2019')]
dYearDelta %>% 
  kable(., caption = "Year-over-Year Rate of Change in Listings by Neighbourhood Group",
        row.names = FALSE)
```

```{r}
ggplot(melt(dYearDelta), aes(x = neighbour_group, y = variable, fill = value)) +
  geom_tile() +
  scale_fill_viridis() +
  ggtitle("Year-over-Year Growth in Listings by Neighbourhood Group")
```

*Average annual growth across all NYC neighbourhood groups was `r mean(as.matrix(dYearDelta[,-1]))`. The neighbourhoods with the strongest growth year-over-year were Bronx, Queens, and Staten Island.*

### Plot: Distribution of Log Listing Price Over Time

```{r}
dfplot2 <- df[,c('year', 'price')]
dfplot2$year <- as.factor(dfplot2$year)
# log transform price
dfplot2$logprice <- log(dfplot2$price)
ggplot(dfplot2, aes(y = logprice, factor(year), fill = year)) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE) +
  xlab('year') +
  ylab('log of listing price') +
  ggtitle("Log Listing Price by Year")
```

*Prices haven't changed too much year-over-year. The average listing price across all listings stayed relatively stable around `r round(mean(dfplot2$price),2)`.*

### Plot: Distribution of Listing Price Over Time by Neighbourhood Group

```{r}
dfplot3 <- df[,c('year', 'neighbourhood_group', 'price')]
dfplot3$year <- as.factor(dfplot3$year)
# log transform price
dfplot3$logprice <- log(dfplot3$price)
ggplot(dfplot3, aes(y = logprice, factor(year), fill = year)) +
  geom_boxplot(color = 'white') +
  scale_fill_viridis(discrete = TRUE) +
  xlab('year') +
  theme_dark() +
  ylab('log of listing price') +
  facet_wrap(~neighbourhood_group) +
  ggtitle("Log Listing Price by Year",
          subtitle = "Facets Neighbourhood Group")
```

*Again, prices appeared to have remained stable over the years even by neighbourhood group.*

# Visualization of Key Variables for Modeling

```{r}
# cleaned listings data for 2019, including census attributes
df <- fread('data/listings_cleaned.csv')
```

### Plot: How Does Review Data Affect Price?

```{r}
keep_cols <- c(names(df)[grepl(pattern = 'review', names(df))], 'price')
dfplot <- as.data.frame(df) %>% .[, keep_cols] %>%
  .[, sapply(., is.numeric)] %>% .[complete.cases(.),]
dfplot <- melt(cor(dfplot))
```

```{r}
dfplot[dfplot$Var1 == 'price',] %>%
  kable(., caption = "Correlation of Review Variables to Price", row.names = FALSE)
```

```{r}
ggplot(dfplot, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Correlation Matrix of Review Variables and Price")
```

*Price is not highly correlated with any of the review variables.It probably means that hosts aren't setting prices based on what people are saying about the rentals and pricing is more influenced by market forces/other factors.*

### Plot: How Does Neighborhood-Based Census Data Correlate with Price?

```{r}
keep_cols <- c(
  'population','population_density','housing_units','occupied_housing_units',
  'median_home_value','median_household_income','price'
)
dfplot <- as.data.frame(df) %>% .[, keep_cols] %>%
  .[, sapply(., is.numeric)] %>% .[complete.cases(.),]
dfplot <- melt(cor(dfplot))
```

```{r}
dfplot[dfplot$Var1 == 'price',] %>%
  kable(., caption = "Correlation of Census Variables to Price", row.names = FALSE)
```

```{r}
ggplot(dfplot, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Correlation Matrix of Census Variables and Price")
```

*Price isn't highly correlated with Census attributes. However, as expected there is some moderate correlation between price, home value, and household income.*

### Plot: How Does Neighbourhood Persona (PRIZM Attributes)  Relate with Price?

```{r}
# discretize price into 5 equal frequency bins
df$priceDisc <- arules::discretize(df$price, method = 'frequency', categories = 5)
```

```{r}
dat <- as.data.frame(df)
# remove prizm_cluster = 'UNK' (unknown)
dat <- dat[!dat$prizm_cluster == 'UNK',]
# keep just the top prizm clusters that make up 80% of the data
clust_freq <- sort(table(dat$prizm_cluster), decreasing = TRUE)
clust_keep <- clust_freq[which(cumsum(clust_freq) <= 0.8 * sum(clust_freq))] %>% names
dat <- dat[dat$prizm_cluster %in% clust_keep,]
# normalize the prizm features for each cluster
# function to normalize vector
normPrizm <- function(x) {
  paste(
    unique(
      do.call('c', strsplit(gsub("\\[|\\]|[\"']",'',x), split = ', '))
    ), collapse = ' | '
  )
}
for (i in clust_keep) {
  dat$prizm_features_norm[dat$prizm_cluster == i] <- normPrizm(
    dat$prizm_features[dat$prizm_cluster == i]
  )
}
# contingency table of clusters and discretized price
dfPlot <- table(dat[,'prizm_cluster'], dat[,'priceDisc']) %>% as.data.frame
# reorder cluster factor levels
dfPlot$Var1 <- factor(dfPlot$Var1, levels = clust_keep)
# scale frequency relative to cluster
for (i in clust_keep) {
  dfPlot$Freq[dfPlot$Var1 == i] <- scale(dfPlot$Freq[dfPlot$Var1 == i])
}

```

```{r}
ggplot(dfPlot, aes(Var1, Var2)) + 
  geom_point(size = 10, aes(color = Freq)) + 
  scale_color_viridis() +
  xlab("Cluster Label") + ylab("Price Bins") +
  ggtitle("Contingency Plot of Discretized Price and Top 14 Neighbourhood Personas")
```

```{r}
desc = sapply(clust_keep, function(x) unique(dat$prizm_features_norm[dat$prizm_cluster == x]))

dfClustDesc <- data.frame(
  cluster = clust_keep,
  description = desc
)
# table of cluster definitions
kable(dfClustDesc, row.names = FALSE)
```

Takeaways:

* Neighbourhood personas that generally price their listings lower are:
    * Low-Rise Living | Urban Elders | Aspiring A-Listers | City Roots | Multi-Culti Mosaic | Connected Bohemians | Multi-Culti Families | Urban Modern Mix
* Neighbourhood personas that generally price their listings higher, but to a lesser proportion include:
    * Young Digerati | Urban Elders | Connected Bohemians | Money & Brains | The Cosmopolitans | Aspiring A-Listers
  